\LARGE \bf Laser Tracker Placement Optimization for highly flexible manufacturing systems

### 2024-07-25 paper/paper.tex
updaeted penalty for minimization and range objective in paper

    1 & \text{if } \text{marker is in line of sight of the laser tracker} \\

@@

    1 & \text{if } \text{marker is in the field of view of the laser tracker} \\

@@

    g(p_m,p_l) = \cos(\alpha)

@@

    h(p_m,p_l) = \lambda f(p_m,p_l) + (1-\lambda) g(p_m,p_l)

@@

@@

The pose of the laser tracker is in reality also defined by a transformation matrix $T_{l}$.
However since the laser tracker has a spherical measurment volume we only need to define the position of the laser tracker $p_{p_l}$, which is a three dimensional vector.
The laser tracker is also mounted on a stand, in principle this can vary in height but only in a limited range.
This can be encoded using a simple constraint:
\begin{equation}
    h_{min} \leq p_{p_l,z} \leq h_{max}
\end{equation}
where $h_{min}$ and $h_{max}$ are the minimum and maximum height of the laser tracker stand.
The laser tracker also has a limited range of measurement.
Here we can define a constraint that the distance between the laser tracker and the marker is within the range of the laser tracker:
\begin{equation}
    d_{min} \leq ||p_{p_m} - p_{p_l}|| \leq d_{max}
\end{equation}
where $d_{min}$ and $d_{max}$ are the minimum and maximum range of the laser tracker.
The final design space is then 9 dimensional, 6 for the marker pose and 3 for the laser tracker pose.

\section{Optimization algorithm}
There are a large number of possible optimization algorithms that can be used to solve this problem.
Following the no free lunch theorem we want to pick the algorithm that makes the most use of the underlying structure of the problem. %TODO citation needed
In this case the problem is a continuous optimization problem with a relatively low number of design variables.
On the other hand it is a gradient free optimization problem due to the line of sight visibility function.
These two properties make the problem well suited for the particle swarm optimization (pso) algorithm.
Particle swarm optimization (PSO) is a computational method for finding the optimal solution to a problem
 It is inspired by the behaviour of swarms in nature, such as flocks of birds or schools of fish, which exhibit emergent behaviour that is intelligent and efficient.
 In PSO, a group of particles (also called agents or individu-als) move through a search space, according to the fol-lowing update rules:
\begin{equation}
        \begin{split}
        x_{i,j} &= x_{i,j} + v_{i,j} \\
    v_{i,j} &= wv_{i,j} + c_1r_1(p_{i,j}-x_{i,j}) + c_2r_2(p_{g,j}-x_{i,j})
        \end{split}
\end{equation}
where $x_{i,j}$ is the position of particle $i$ in dimension $j$, $v_{i,j}$ is the velocity of particle $i$ in dimension $j$, $w$ is the inertia weight,
 $c_1$ and $c_2$ are the cognitive and social coefficients, $r_1$ and $r_2$ are random numbers between 0 and 1, $p_{i,j}$ is the best position of particle $i$ in dimension $j$ and $p_{g,j}$
is the best position of the swarm in dimension $j$.
In our case the particles are a 9 dimensional vector based on the position of the laser tracker and the pose of the marker.
The best particle is the one that maximizes the objective function defined in equation \ref{eq:objective}.
The hyperparameters we need to tune are collected in table \ref{tab:hyperparameters}.
Note that the pso algorithm has no direct way to encode the constraints we defined earlier.
Instead we can use a penalty function that penalizes any particles that violate the constraints.
This is a common approach in optimization and is also used in the implementation of the pso algorithm in the scipy library. %TODO citation needed
The penalty function can be defined as:
\begin{equation}
    p(x) =  \begin{cases}
        0 & \text{if } x \text{ satisfies all constraints} \\
        -\rho & \text{otherwise}
    \end{cases}
\end{equation}
where $\rho$ is a large penalty factor.

>>>>

    -1 & \text{if } \text{marker is in line of sight of the laser tracker} \\

@@

    -1 & \text{if } \text{marker is in the field of view of the laser tracker} \\

@@

    g(p_m,p_l) = -\cos(\alpha)

@@

    h(p_m,p_l) = \int \lambda f(p_m,p_l) + (1-\lambda) g(p_m,p_l) dt

@@

The intergral is taken over the entire trajectory of the manufacturing system that needs to be observed by the laser tracker.

@@

The pose of the laser tracker is in reality also defined by a transformation matrix $T_{l}$.
However since the laser tracker has a spherical measurment volume we only need to define the position of the laser tracker $p_{p_l}$, which is a three dimensional vector.
The laser tracker is also mounted on a stand, in principle this can vary in height but only in a limited range.
This can be encoded using a simple constraint:
\begin{equation}
    h_{min} \leq p_{p_l,z} \leq h_{max}
\end{equation}
where $h_{min}$ and $h_{max}$ are the minimum and maximum height of the laser tracker stand.
The laser tracker also has a limited range of measurement.
Here we could define a constraint that the marker needs to be within the measurement range of the laser tracker.
But in practice the laser tracker works better the closer the marker is to the laser tracker so long it is in the minimum range.
This can be encoded using a penalty function:
\begin{equation}
    p(p_m,p_l) =  \begin{cases}
        d & \text{if } \text{marker is in measurement range of the laser tracker} \\
        \rho & \text{otherwise}
    \end{cases}
\end{equation}
where $d$ is the distance of the marker to the laser tracker and $\rho$ is a large penalty factor.
The final optimization problem can be formulated as follows:
\begin{equation}
    \begin{split}
    \max_{p_m,p_l} & h(p_m,p_l) \\
    \text{s.t.} & \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} \leq r \\
    & 0 \leq p_{p_m,z} \leq h \\
    & h_{min} \leq p_{p_l,z} \leq h_{max} \\
    & p(p_m,p_l) \geq 0
    \end{split}
\end{equation}

\section{Optimization algorithm}
There are a large number of possible optimization algorithms that can be used to solve this problem.
Following the no free lunch theorem we want to pick the algorithm that makes the most use of the underlying structure of the problem. %TODO citation needed
In this case the problem is a continuous optimization problem with a relatively low number of design variables.
On the other hand it is a gradient free optimization problem due to the line of sight visibility function.
These two properties make the problem well suited for the particle swarm optimization (pso) algorithm.
Particle swarm optimization (PSO) is a computational method for finding the optimal solution to a problem
 It is inspired by the behaviour of swarms in nature, such as flocks of birds or schools of fish, which exhibit emergent behaviour that is intelligent and efficient.
 In PSO, a group of particles (also called agents or individu-als) move through a search space, according to the fol-lowing update rules:
\begin{equation}
        \begin{split}
        x_{i,j} &= x_{i,j} + v_{i,j} \\
    v_{i,j} &= wv_{i,j} + c_1r_1(p_{i,j}-x_{i,j}) + c_2r_2(p_{g,j}-x_{i,j})
        \end{split}
\end{equation}
where $x_{i,j}$ is the position of particle $i$ in dimension $j$, $v_{i,j}$ is the velocity of particle $i$ in dimension $j$, $w$ is the inertia weight,
 $c_1$ and $c_2$ are the cognitive and social coefficients, $r_1$ and $r_2$ are random numbers between 0 and 1, $p_{i,j}$ is the best position of particle $i$ in dimension $j$ and $p_{g,j}$
is the best position of the swarm in dimension $j$.
In our case the particles are a 9 dimensional vector based on the position of the laser tracker and the pose of the marker.
The best particle is the one that maximizes the objective function defined in equation \ref{eq:objective}.
The hyperparameters we need to tune are collected in table \ref{tab:hyperparameters}.
Note that the pso algorithm has no direct way to encode the constraints we defined earlier.
Instead we can use a penalty function that penalizes any particles that violate the constraints.
This is a common approach in optimization and is also used in the implementation of the pso algorithm in the scipy library. %TODO citation needed
The penalty function can be defined as:
\begin{equation}
    p(x) =  \begin{cases}
        0 & \text{if } x \text{ satisfies all constraints} \\
        \rho & \text{otherwise}
    \end{cases}
\end{equation}
where $\rho$ is a large penalty factor.
